---
title: 'Assignment 3: Vis'
author: "Sage Grey"
date: "9/27/2020"
output: html_document
---

#### **Assignment Overview:**
Creative Assignment 3: Areas and locations (points/polygons) (due before class on September 28)

Identify a municipal open data portal and find two point layers (A and B) and two polygon layers (C and
D), where the polygons do not cover the entire municipality (in other words, you shouldn’t use the set of all
census tracts as one of your layers, but you could use a subset of census tracts that have particular
characteristics). 

Calculate eight of the following 15 possible metrics, and illustrate the results of each calculation with a
map.
1. The number and proportion of A points within a specified distance of B points.
2. The number and proportion of B points with a specified distance of A points.
3. The average (Euclidean) distance between A points and B points.
4. The average (Euclidean) distance between A points and their nearest respective B points.
5. The average (Euclidean) distance between B points and their nearest respective A points.
6. The number and proportion of A points within C polygons.
7. The number and proportion of A points within D polygons.
8. The number and proportion of B points within C polygons.
9. The number and proportion of B points within D polygons.
10. The number and proportion of C polygons containing A points.
11. The number and proportion of D polygons containing A points.
12. The number and proportion of C polygons containing B points.
13. The number and proportion of D polygons containing B points.
14. The number and proportion of C polygons that overlap with D polygons.
15. The number and proportion of D polygons that overlap with C polygons.


```{r setup, include=FALSE}
library(sf)
library(tidyverse)
library(ggthemes)
library(ggspatial)
library(units)
library(nngeo)
library(sf)
```

```{r}
nhoods_bos <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/3525b0ee6e6b427f9aab5d0a1d0a1a28_0.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D", quiet = TRUE )

pol_dist_bos <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/9a3a8c427add450eaf45a470245680fc_5.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D",quiet =TRUE)

hosp_bos <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/8f1e5741a4ee4c0f82925d17aab3c002_2.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D", quiet =TRUE)

bos_police_stations <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/e5a0066d38ac4e2abbc7918197a4f6af_6.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D")

non_public_schools <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/0046426a3e4340a6b025ad52b41be70a_1.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D")

public_schools <-st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/1d9509a8b2fd485d9ad471ba2fdb1f90_0.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D")

public_libraries <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/cb00f9248aa6404ab741071ca3806c0e_6.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D")

soc_vulnerab <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/34f2c48b670d4b43a617b1540f20efe3_0.kml?outSR=%7B%22latestWkid%22%3A3857%2C%22wkid%22%3A102100%7D")

bos_zip_code <-  st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/53ea466a189b4f43b3dfb7b38fa7f3b6_1.geojson?outSR={%22latestWkid%22:2249,%22wkid%22:102686}")

fio_records <- st_read("https://data.boston.gov/dataset/4ebae674-28c1-4b9b-adc3-c04c99234a68/resource/03f33240-47c1-46f2-87ae-bcdabec092ad/download/mark43_fieldcontacts_for_public_20192.csv")
```


#SETUP Coordinate System
```{r}
MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"

nhoods_bos <- nhoods_bos %>%
  st_transform(MA_state_plane)

pol_dist_bos <- pol_dist_bos %>%
  st_transform(MA_state_plane)

hosp_bos <- hosp_bos %>%
  st_transform(MA_state_plane)

bos_police_stations <- bos_police_stations %>%
  st_transform(MA_state_plane)

non_public_schools <- non_public_schools %>%
  st_transform(MA_state_plane)

bos_zip_code <- bos_zip_code %>%
  st_transform(MA_state_plane)

public_schools <- public_schools%>%
  st_transform(MA_state_plane)

public_libraries <- public_libraries %>%
  st_transform(MA_state_plane)

soc_vulnerab <- soc_vulnerab %>%
  st_transform(MA_state_plane)
```



# Setup Map
```{r}
ggplot(nhoods_bos) +
 # geom_sf(fill = NA, color ="black") +
  geom_sf(data=hosp_bos, color ="sienna", size = 2, alpha =.75) +
  geom_sf(data=pol_dist_bos, fill =NA, color ="gray") +
  geom_sf(data =bos_zip_code, fill ="lightblue", color="darkgreen", alpha =.25) +
  geom_sf(data =public_schools, color = "rosybrown4") +
  theme_map() +
  annotation_scale()
```


## Creating A Buffer (200 meters from a public school)
```{r}
school_buffer <- st_buffer(public_schools, dist =200) %>% 
  st_union()

ggplot(school_buffer) +
  geom_sf(data = nhoods_bos, fill =NA) +
  geom_sf(color = "paleturquoise3", size=1.4 ) +
  #geom_sf(data =public_libraries, color ="orangered2") +
  geom_sf(data =non_public_schools, color = "royalblue3") +
  theme_map()
```


===
### Subsetting Points with a polygon 
```{r}
public_private_schools <- non_public_schools[school_buffer,]

ggplot(school_buffer) +
  # geom_sf(data = nhoods_bos, fill =NA) +
  geom_sf() +
  geom_sf(data =public_private_schools,
          color = "royalblue3",
          size = 2) +
  theme_map()
```

```{r}
non_public_schools <- non_public_schools %>% 
  st_join(public_private_schools) %>%
  mutate(by_public = !is.na(Name.y))

n_private_public <-sum(non_public_schools$by_public)

n_private_public
```
```{r}
n_non_public_schools <- length(non_public_schools$by_public)

pct_PPP <- n_private_public / n_non_public_schools 

pct_PPP
```
```{r}
left_side <- st_bbox(non_public_schools)$xmin
right_side <- st_bbox(non_public_schools)$xmax
top_side <- st_bbox(non_public_schools)$ymin


ggplot(nhoods_bos) +
  geom_sf() +
  geom_sf(data = non_public_schools, aes(color = by_public)) +
  scale_color_manual(values =c("#94A89A", "#A44A3F"),
                     name ="Private Schools \n by distance to a public school",
                     labels = c("No Public School within 200 Meters",
                                "Public School within 200 Meters")) +
  
  #annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text",
           x = right_side, 
           y = top_side, 
           label = paste("Of the ", 
                         prettyNum(n_non_public_schools, big.mark = ","),
                         " private \nschools in Boston\n", 
                         prettyNum(n_private_public, big.mark = ","),
                         " (", 
                         prettyNum(100*pct_PPP, digits = 0),
                         "%) are within 200\nmeters of a public school.",
                         sep = ""),
           hjust = 0, vjust = 0, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "#D4E09B"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "#D4E09B"))
  

```




#### Chloropleth FIO Records by ZipCodes
```{r}
ggplot(fio_records, aes(x=zip)) +
   theme(axis.text.x = element_text(angle=65, vjust=0.6)) +
  coord_flip() +
  geom_bar()
```
```{r}


freq_stop <- as.data.frame(table(unlist(fio_records$zip))) %>%


zip_stops <- bos_zip_code %>%
  left_join(freq_stop, bos_zip_code %>%as.data.frame(), by =c("Var1" = "ZIPS"))%>%
  select(ZIPS, Freq, OBJECTID,geometry)


```


ggplot(zip_stops) +
  geom_sf(data=freq_stop,
          aes(fill = Freq) +
  scale_fill_viridis_c(name = "Blah Blah Blah",
                       breaks = breaks <- seq(0, 30000, by = 5000),
                       labels = paste(prettyNum(breaks, big.mark = ","))) +
  theme_map() 
```



